---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup}
#| include: false
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## Purpose

<!-- badges: start -->
<!-- badges: end -->

This repository contains an R package with a focus on Mixture of Experts (MoE)
    applied to Bayesian linear and censored regression.



## Installation

You can install the development version of `bmoe` from [GitHub](https://github.com/) with:

```{.r}
devtools::install_github("nclJoshCowley/bmoe", auth_token = ...)
```



## Model Description

TODO: Include model description here.



## Worked Example

### Simulation

We can simulate data from these models using `bmoe::simulate_bmoe()`.

Alternatively, one can use the example wrapper to utilise default arguments.

```{r}
example_sim <- bmoe::example_simulate_bmoe()

example_sim$data
```

### Fitting

We require prior hyperparameters for all models; the default is a vague prior
for each parameter but the number of components ($K$) is assumed known and must
be set by the user.

```{r}
example_prior <- bmoe::bmoe_prior(k = 3)

example_prior
```

For simulation study results, one can pass a simulation directly.

```{.r}
bmoe::bmoe(example_sim, prior = example_prior)
```

More generally, this package provides a extended formula-data interface.

```{.r}
example_fit <-
  bmoe::bmoe(
    y01 ~ x01 + x02 + x03,
    data = example_sim$data,
    prior = example_prior
  )
```

This interface goes beyond the base R formula system as we can

* model multiple response variables as conditionally independent
  (conditional on component membership) use the `+` symbol in the LHS.
  
* allow two sets of predictors can be separated for regression purposes and
    component probability weighting purposes using `|`.
    
For example, `y01 + y02 ~ x01 + x02 + x03 | x03` implies two response
    variables, `y01` and `y02`, to be regressed against the linear
    predictor formed from `x01 + x02 + x03` according to some component
    probabilities based on the linear predictor formed from `x03`.

### Reporting

Analysis reports can be generated from any fitted object and desired filename.

```{.r}
bmoe::render_bmoe_fit(example_fit, "report-name")
```



## Future Work

Extend this `README` with more functionality including relabelling, prediction
and log likelihood.
