---
title: "Memoisation of Component Allocation Draws"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{component-memoisation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup}
#| include: false
library(bmoe)
library(knitr)

knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```

## Purpose

Explains the reason for [memoisation](https://memoise.r-lib.org/) within key
    functionality and nuance that should be understood by the users of this
    package.


## Motivation

Suppose we fit a Bayesian Mixture of Experts analysis.

```{.r}
object <- example_bmoe_fit(multiple_y = TRUE)
new_data <- object$new_data[1:4, ]
```

And with the `new_data`, defined above, we want to 

1. calculate the OOS log likelihood per observation;

1. make predictions based on the predictors in the new data.

The functionality to achieve these objects are isolated in

1. `bmoe::extract_pointwise_log_lik()`;

1. `bmoe:::predict.bmoe_fit()` (callable by `stats::predict`).

Both of these functions use

```{.r}
z <- calculate_component_samples(object, new_data)
```

to draw new samples from the posterior distribution.

While the distribution will be the same when all inputs are the same, the
    samples may be slightly different due to the randomness of the quantity.
    
Hence, we memoise the function to extract samples so the draws denoting the
    component allocations remain unchanged unless the `object` or `new_data`
    are changed.

To see this, consider the following cleaned output from the R console.

```{.r}
memoise::is.memoised(calculate_component_samples)
#> [1] TRUE


system.time({
  z <- calculate_component_samples(object, new_data)
  print(z)
})
#> [MESSAGE]: Drawing new allocation samples from relevant distribution
#>
#> Output summarised over 10000 iterations and 2 chains:
#> 
#> varname = 'z'
#> 
#> [1] 2.08050 2.00210 2.17815 1.83230
#>    user  system elapsed 
#>    5.59    0.16    5.73


system.time({
  z <- calculate_component_samples(object, new_data)
  print(z)
})
#> Output summarised over 10000 iterations and 2 chains:
#> 
#> varname = 'z'
#> 
#> [1] 2.08050 2.00210 2.17815 1.83230
#>    user  system elapsed 
#>    0.02    0.00    0.02


system.time({
  z <- calculate_component_samples(object, new_data[1:2, ])
  print(z)
})
#> [MESSAGE]: Drawing new allocation samples from relevant distribution
#>
#> Output summarised over 10000 iterations and 2 chains:
#>
#> varname = 'z'
#> 
#> [1] 2.0798 1.9923
#>    user  system elapsed 
#>    5.02    0.12    5.13
```

```{r}
#| include: false
if (interactive()) {
  object <- example_bmoe_fit(multiple_y = TRUE)
  new_data <- object$new_data[1:4, ]

  # Source code for above 'stored' output
  memoise::is.memoised(calculate_component_samples)
  
  system.time({
    z <- calculate_component_samples(object, new_data)
    print(z)
  })
  
  system.time({
    z <- calculate_component_samples(object, new_data)
    print(z)
  })
  
  system.time({
    z <- calculate_component_samples(object, new_data[1:2, ])
    print(z)
  })
}
```


We observe that the second attempt to draw `z` from the new data is returned
    almost immediately, with the same values.
Whereas modifying the inputs forces the cache to be invalid and new draws are
    produced.
    
## Conclusion

Using the same `object` and `new_data` in the pointwise log likelihood
    function and prediction function will use the same component allocation
    samples.

Note that this functionality will not persist across R sessions and so a
    message is produced when `calculate_component_samples` samples new draws
    from the posterior.
